### 2011-05-07 paper/paper.tex

\documentclass{article}

% required includes
\usepackage{graphicx}
\usepackage{url}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
% styles for flowcharts
\tikzstyle{machine} = [star, star points=7, star point ratio=0.8, draw, text width=5em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{data} = [trapezium, trapezium left angle=70, trapezium right angle=-70, draw,text centered]
\tikzstyle{process} = [rectangle, draw, text width=5em, text centered, rounded corners, minimum height=4em]

\addtolength{\textwidth}{10pt}

\title{CAT-DWAM, A Question Answering System}
\author{
Alec Story,
Ansu Abraham,
Craig Frey,\\
Dustin Tiedemann,
Michael Zhu,
Thomas Levine,
Whitney Foster\\
}

\begin{document}
\maketitle

\section{Overview}

% What are you doing. Why are you doing it.
% ``this is effectively a statement of your hypothesis, so you should
% include justification of why you think that the particular approach
% will be effective

\section{The QA System}
% we need:
% a detailed walkthrough of what our system did to handle one question
% (in general) in the corpus. include the system output (enough to
% convince them that it is actually doing what we say)

% this is pretty much the final system proposal from part
% one. delete/rewrite if necessary (we need a description of the
% baseline in our final report)
We implemented our question answering system in Python, and use external
tools such as the natural language toolkit.

Our design mimics the design of Watson, with the intention of trying out a
variety of techniques to select answers from documents, and to use a machine
learning system to choose the appropriate one.  See figure \ref{diagram} for
a graphical overview.

We first shred the documents for each question into all of the component verb
and noun phrases.  We pass these candidates into a battery of evaluators that
assign a confidence value to each of them, and pass these confidences into a
machine learning algorithm that combines them to produce a ranked output.

From this ranked output we choose our final answers.

\begin{figure}
 \label{diagram}
 \begin{center}
  \begin{tikzpicture}[node distance=2.5cm, auto, >=stealth]
   % nodes
   \node[data] (Q)                                      {Question};
   \node[process] (ch) [below of=Q]                     {Chunker};
   \node[data] (docs) [right of=ch]                     {Documents};
   \node[process] (fil) [below of=ch]                     {Filter};
   \node[process] (a2) [below of=fil]                    {Answer Evaluator};
   \node[process] (a1) [left of=a2]                     {Answer Evaluator};
   \node[process] (a3) [right of=a2]                    {Answer Evaluator};
   \node[machine] (as) [below of=a2]                    {Score aggregator};
   \node[data] (A) [below of=as]                        {Answer};

   % edges
   	\draw[->] (Q) to (ch);
	\draw[->] (docs) to (ch);
	\draw[->] (ch) to (fil);
	\draw[->] (fil.south) to [out=270, in=90] (a1);
	\draw[->] (fil.south) to [out=270, in=90] (a2);
	\draw[->] (fil.south) to [out=270, in=90] (a3);
	\draw[->] (Q.west) to [out=180, in=180] (a1);
	\draw[->] (Q.west) to [out=180, in=180] (a2);
	\draw[->] (Q.west) to [out=180, in=180] (a3);
	\draw[->] (a1.south) to [out=270, in=135] (as);
	\draw[->] (a2.south) to [out=270, in=90] (as);
	\draw[->] (a3.south) to [out=270, in=45] (as);
	\draw[->] (as) to (A);
  \end{tikzpicture}
  \caption{Data Flow}
  \label{flowchart}
 \end{center}
\end{figure}

\subsection{Parsing the Corpus Documents}
% Talk about Whitney's Shredder module here
	To get candidate answers we took in the question number, parsed the
	documents associated with it by chunking it into syntactically correlated
	parts of words (noun phrases, verb phrase, etc), and outputted a list of
	tuple candidate answers.  These were in the format (string of candidate
	answer, document number, word index in the document, the phrase tag,
	question number).

	To chunk the documents for the question, we prepared them by separating out
	the punctuation and the words with a space so that the taggers would work.
	We then POS tag all the words in the document and then get the IOB chunk
	tags for all the words. This data is then run through the conll.2000 chunker
	(http://www.clips.ua.ac.be/conll2000/chunking/) that labels what type of
	phrase the word belongs to (noun phrase, verb phrase, etc). The words in
	the same phrase are then combined together to form one long phrase as the
	candidate answer string.

\subsection{Question Evaluators}

\subsubsection{Question-Document Comparison}

	We implemented a system to compare the question and answer against the
	proposed document in two ways:  one, with a na\"ive diffing algorithm that
	simply seeks out the longest indentical sequence, and one by implementing
	the Smith-Waterman local alignment
	algorithm.\footnote{\url{http://en.wikipedia.org/wiki/Smith-Waterman_algorithm}}
	We also implemented a system to attempt to re-write questions, discussed
	below, and applied these same techniques to the re-written question

	The na\"ive method does an adequate job when the document highly resembles
	the question, but does very poorly otherwise because it has no tolerance of
	faults.  The local alignment does much better for documents that resemble
	the question a great deal, but are not precise - it can smoothly deal with
	small differences, and still produce a good alignment.  This is a standard
	technique from bioinformatics, and is used extensively to align small
	sequences of DNA, RNA or protein against a larger corpus, such as a
	chromosome or a whole genome.  This is a fairly similar usage to aligning a
	question against a document which is much longer than the question.

	We chose Smith-Waterman over the similar Needleman-Wunsch, of which is a
	special case, because Needleman-Wunsch creates a \emph{global} alignment,
	which forces all of the characters in both strings to align.  This leads to
	the insertion of many gaps, and needlessly aligns words which are useless,
	for example, ``Who is Queen of England?'' should align well against
	``Elizabeth is Queen of England'' and it does, under Smith-Waterman, but the
	``who'' causes Needleman-Wunsch to choke.  Similarly, Needleman-Wunsch does
	not excuse one string for being shorter than the other, and chokes on all of
	the apparent gaps that it needs to insert, so it is not a good choice.

	We locate what we best believe to be the question via these two methods, and
	then evaluate how far away from that segment of the document the answer is.
	If it is fairly close, that is a good indicator that it is highly related to
	the question, but if it is far, it is likely to be unrelated.  We output
	that distance, and either the length of the na\"ive alignment or the score
	that Smith-Waterman produces (higher is better for both) because very short
	or very poor alignments are less likely to reflect an actual representation
	of the question.

\subsubsection{Punctuation}

	1 if the candidate answer is immmediately followed by a comma, period,
	quotation marks, semicolon, or exclamation mark, and 0 otherwise. This is a
	good heuristic by which to rank candidate answers since it retrieves a
	property from the context where the answer appears in the document.

\subsubsection{Apposition Filter}
% Alec

	Because we thought it would be difficult to analyze complete parse
	information for questions, we decided to tackle apposition by a simple
	heuristic:  does "question, answer," align better than "question answer"?
	If it does, then the answer is likely to be in apposition.  We also applied
	this to re-written questions.

\subsubsection{Number and Quantity Answer Type Filters}
% Tom
We used regular expressions to identify answer candidates that contained digits.
We classified these candidates as number-type answers or quantity-type answers.
Answer candidates containing exclusively digits were classified as number-type
answers, and answer candidates containing digits and other characters were classified
as quantity-type answers.

\subsubsection{Measurement Units or Matching Type Filters}
% Tom
For quantity-type answer candidates, we used simple regular expressions to extract
the proceeding and preceeding space-delimited phrases. This is quite close to a
lay definition of ``word''; for the answer candidate ``about 350 km/h'', the
phrases ``about'' and ``km/h'' would be extracted as units. We extracted the units
from the training corpus.

For test answers, we searched the context of the answer for colocated units
matching what we found in the training corpus. We returned a score of 1 if there
was a unit and a score of 0 if there wasn't.
\subsubsection{Sequence Length}
% Mike

\subsubsection{Word Vector Comparison}

This evaluation measures the angle difference between the word vector of the
question and the answer, after removing stop words, and scaled to [0,1].

\subsubsection{Bag of Words Filter}

Simply how many times the words in the question (without stop words) appear in
the answer, a simpler but less precise metric than the word vector.

\subsubsection{Novelty Factor}

Novelty is whether at least one non-stop word in the answer that is not found in
the question.  This is intuitive because an answer that is completely non-novel
is unlikely to be interesting.

We implemented this in two slightly different ways:  one, that just reports the
presence or absence of novel words, and one that counts the number of novel
words in the answer.  We split them this way to make it clear to the learning
algorithm that any novelty is very important, but extra novelty probably doesn't
matter much.

\subsubsection{Part of Speech Matching}
% Craig
The chunker module assigns one of four part of speech tags to each answer candidate
phrase. We discard (assign 0 to) the answer candidate if the tag is either PP (Prepositional Phrase) 
or S (simple declarative clause). If the answer candidate is NP (Noun Phrase) we 
assign it a probability of 1 (most likely). If the answer candidate is VP (Verb Phrase
we assign it a probability of 0.1 (not very likely). It is possible that a higher score would be
obtained if we also discarded verb phrases entirely, since the short question format typically
will point to a noun phrase.

\subsection{Score Aggregation}
We used supervised classification algorithms to aggregate the scores produced
by the separate answer evaluators. We trained the classifiers to predict whether
a candidate answer was correct based on the scores produced by the answer evaluators.

Each of the answer evaluators produced at least one numerical for each answer
candidate; we used these as predictors for the classification algorithms.
If we thought that some of these predictors might interact, we used their products
as predictors as well.

For each candidate answer, we checked whether that candidate was a correct answer based on
the regular expressions. We used the correctness of the answer (correct or incorrect)
as the response. We were thus able to predict whether an answer was correct
based on the scores from the answer evaluators.

The classifiers that we attempted were
\begin{itemize}
\item support vector machines (SVM),
\item Fisher Discriminant Analysis (FDA),
\item Spectral Regression Discriminant Analysis (SRDA) and
\item Penalized Discriminant Analysis (PDA)
\end{itemize}

\subsection{Candidate Thinning}

Our chunker generates over 20,000 candidate answers per question.  This is too
many to efficiently process, so we require the following to be true before we
even attempt to process it:
\begin{itemize}
\item the candidate must be a noun phrase.  Other types of answers were
exceedingly rare in the training data.
\item the answer must be within 50 bytes of the aligned question in the
document.  This is already a feature we train on, but this reduces the volume of
work to process.
\end{itemize}

\subsection{Caching}

Many of our subsystems took a significant amount of time to compute.  In order
to mitigate this, we implemented aggressive caching systems, but they did not
reduce the time to compute too dramatically, and ultimately, the time constraint
was a very serious one.

%We tried these classifiers separately rather than 
We ended using just the SRDA classifier since the others were throwing memory errors.
\section{Results}

% we need:
% the output file of answers produced by our system for the questions
% from the development corpus
% an evaluation using the Mean Reciprocal Rank Evaulation Measure)
% an analysis of the system's performance on questions of the
% development corpus
   % what worked?
   % what didn't work?
   % which component was strongest/weakest?
   % how does our system compare to the baseline?
% responses from our system for test set released on Wednesday
   % we also need to submit this file

\subsection{Baseline Results}
Our baseline system simply takes the first answer of the training data and
repeats that as the answer to all following questions. It is included as
zero.py, and its mean reciprocal rank over 197 questions is 0.048.

This represent a total bare-bones approach with essentially no hope of getting a
question right.

\subsection{Graphical Methods}

\subsection{Best Run Results}

\section{Conclusion}

\end{document}

### 2011-05-07 paper/paper.tex
edited paper.tex

@@
 
+We did pre-process some of the data by running the chunker (which creates
+candidate phrases along with part of speech tags) and creating files in
+groups of ten questions each with all of the candidates. Running the chunker
+took about two minutes per question, so by doing this we were able to have
+the whole set of potential candidates instantly ready to go.
+
 %We tried these classifiers separately rather than 

### 2011-05-07 paper/paper.tex

@@
 \begin{itemize}
-\item support vector machines (SVM),
+\item Support Vector Machines (SVM),
 \item Fisher Discriminant Analysis (FDA),
@@
 %We tried these classifiers separately rather than 
-We ended using just the SRDA classifier since the others were throwing memory errors.
+%We ended using just the SRDA classifier since the others were throwing memory errors.
+The implementation of these classifiers in mlpy required too much memory for
+us to run the models on the all of the candidate answers. We used two main approaches
+to solve these problems.
+\begin{itemize}
+\item We used SRDA because it seemed to handle the largest sample size.
+\item We created a decision list from some of our answer evaluators. We ran the anwers
+through this decision list before sending them to the SRDA learning algorithm.
+\end{itemize}
+
 \section{Results}

### 2011-05-07 one.py
?

@@ -99,7 +99,6 @@ def main():
     [punc_loc]
 #     bag_of_words,
 #     novelty_bool]
-#    [pos_test]
 #    [seq_length,punc_loc,question_apposition,rewrite_apposition,pos_test,vector_bag,bag_of_words,novelty_bool] #,novelty_count]
 #    [novelty_count]
     ]
@@ -117,4 +116,4 @@ if __name__ == '__main__':
     main()
     align.load_cache()
     align.save_cache()
-    question_candidates (243)
+#    question_candidates (243)

### 2011-05-07 question_type.py

@@ -32,5 +32,5 @@ def main():
   return answer_type(read_questions.read_questions_answers())
 
 if __name__ == "__main__":
-  print get_question_types()
-  print classify_questions()
+#  print get_question_types()
+  print classify_questions()['Where']

### 2011-05-07 paper/paper.tex
paper

@@
 was a unit and a score of 0 if there wasn't.
+
 \subsubsection{Sequence Length}
 % Mike
+This evaluation measures the difference between sequence length of the question
+and the answer. It finds the length of the maximum sequence between the
+question and answer and also finds the length of the maximum sequence between
+the rewritten question and answer. A sequence is a list of (not necessarily 
+consecutive) words that are shared by two sentences.
 
@@
 
-We did pre-process some of the data by running the chunker (which creates
-candidate phrases along with part of speech tags) and creating files in
-groups of ten questions each with all of the candidates. Running the chunker
-took about two minutes per question, so by doing this we were able to have
-the whole set of potential candidates instantly ready to go.
-
 %We tried these classifiers separately rather than 

### 2011-05-07 one.py
fix lack of return

@@ -87,22 +87,21 @@ def writeAnswers(stuff,filename='tmp-answers.txt'):
 
 def main():
     foo=map(int,question_type.classify_questions(1)['Where'])
-    trainIDs=foo[:-3]
-    validationIDs=foo[-3:]
-#    print trainIDs
-#    trainIDs=range(201,220)
-#    validationIDs=range(339,339)
-#    testIDs=range(338,338)
+    evaluatorCombinationID=10
+#    trainIDs=foo[:-3]
+#    validationIDs=foo[-3:]
+    trainIDs=foo[:4]
+    validationIDs=foo[4:5]
     evaluator_combinations=[
-#    [],
-#     [seq_length,
-    [punc_loc]
-#     bag_of_words,
-#     novelty_bool]
-#    [seq_length,punc_loc,question_apposition,rewrite_apposition,pos_test,vector_bag,bag_of_words,novelty_bool] #,novelty_count]
-#    [novelty_count]
+    [punc_loc,novelty_bool],
+    [punc_loc],
+    [novelty_bool]
+
+#Don't use because they're weird
+#    vector_bag,
+#     [seq_length]
+#    [punc_loc]
     ]
-    evaluatorCombinationID=1
     for evaluators in evaluator_combinations:
         y_train,x_train = question_learning_data(evaluators,trainIDs)
 #        print y_train

### 2011-05-07 sequence_length.py

@@ -14,9 +14,9 @@ def seq_length (question, (answer, doc_num, index, features,q_id)):
        
        returns [max_seq_len, max_seq_len_rewrite] (int list)
     """
-    ans = (seq_length_literal (question, (answer, doc_num, index, features,q_id)),
-           seq_length_rewrite (question, (answer, doc_num, index, features,q_id)))
-    return [ans]
+    ans = [seq_length_literal (question, (answer, doc_num, index, features,q_id)),
+           seq_length_rewrite (question, (answer, doc_num, index, features,q_id))]
+    return ans
 
 def seq_length_literal (question, (answer, doc_num, index, features,q_id)):
     """Takes a question and calculates the maximum sequence length

### 2011-05-07 paper/paper.tex
edited paper.tex
paper
merge

--- 2011-05-07 paper/paper.tex
@@
 
-\subsection{Question Evaluators}
+\subsection{Answer Evaluators}
 
@@
 
+\subsection{Other choices}
+Since the system was taking too long to run we made several choices in order
+to get meaningful results on which to report on. We decided to specifically analyze
+the "Where" questions since there were 28 of them, which seemed like an appropriate
+amount on which to train on.
+
 \section{Results}

### 2011-05-07 one.py

@@ -86,6 +86,7 @@ def writeAnswers(stuff,filename='tmp-answers.txt'):
         answersHandle.close()
 
 def main():
+    align.load_cache()
     foo=map(int,question_type.classify_questions(1)['Where'])
     evaluatorCombinationID=10
 #    trainIDs=foo[:-3]
@@ -109,10 +110,9 @@ def main():
         results=run_question_predictions(evaluators,trained,validationIDs)
         writeAnswers(answerFile(results),'results/combination'+str(evaluatorCombinationID)+'.txt')
         evaluatorCombinationID=evaluatorCombinationID+1
-    
-if __name__ == '__main__':
-    align.load_cache()
-    main()
     align.load_cache()
     align.save_cache()
-#    question_candidates (243)
+    
+if __name__ == '__main__':
+    main()
+    question_candidates (243)

### 2011-05-07 paper/paper.tex
edited paper.tex

@@
 
+To speed alignment (which runs in $O(mn)$ where m and n are the length of the strings being aligned, and the documents are several thousand characters), we cached previous alignments, and looked them up by the first 50 bytes of the two strings being aligned.
+We also stored this in a file so that we could use the pre-computed values between runs.
+
 We did pre-process some of the data by running the chunker (which creates
@@
 the "Where" questions since there were 28 of them, which seemed like an appropriate
-amount on which to train on.
+amount which to train on.
 

### 2011-05-07 one.py
where questions ignoring weird ones

@@ -87,12 +87,15 @@ def writeAnswers(stuff,filename='tmp-answers.txt'):
 
 def main():
     align.load_cache()
-    foo=map(int,question_type.classify_questions(1)['Where'])
-    evaluatorCombinationID=10
-#    trainIDs=foo[:-3]
-#    validationIDs=foo[-3:]
-    trainIDs=foo[:4]
-    validationIDs=foo[4:5]
+    #foo=map(int,question_type.classify_questions(1)['Where'])
+    #Where questions
+    foo=[202, 211, 223, 226, 227, 243, 245, 249, 258, 266, 272, 283, 304, 306, 310, 317, 318, 320, 356, 359, 368, 369, 373, 385, 390, 393]
+    evaluatorCombinationID=20
+    trainIDs=foo[:-6]
+    validationIDs=foo[-6:]
+#    trainIDs=foo
+#    validationIDs=[10025,10026]
+
     evaluator_combinations=[
     [punc_loc,novelty_bool],
     [punc_loc],

### 2011-05-07 paper/paper.tex
paper

--- 2011-05-07 paper/paper.tex
@@
 
+\subsection{Walk through of one question}
+- Initial setup of the Q/A system
+    - We collect all the documents that may contain the answer for all qids (20 documents per qid)
+    - We then run the chunker on the document text creating answer candidate phrases
+    - We save the chunker results to a set of files
+- We are given a question id
+- We retrieve the question text from the provided input file
+- We read the chunker results 
+    - We filter out all candidate answers that are not noun phrases
+    - We also filter out candidate answers not within 50 bytes of the aligned question in the document
+- We feed the filtered answers through our answer evaluators specifically
+    - punctuation_location,
+    - novelty_factor
+    - a combination of both
+- The scores of the evaluators are fed into the supervised classification algorithm:
+    - Spectral Regression Discriminant Analysis 
+    - this gives us a confidence measure
+- Sort by confidence measure 
+- In descending order by confidence measure, pack them into 50-byte answers which are our final outputs
+- The answers are then passed into our scorer
+
 \subsection{Graphical Methods}

### 2011-05-07 one.py
oasethusao

@@ -89,12 +89,12 @@ def main():
     align.load_cache()
     #foo=map(int,question_type.classify_questions(1)['Where'])
     #Where questions
-    foo=[202, 211, 223, 226, 227, 243, 245, 249, 258, 266, 272, 283, 304, 306, 310, 317, 318, 320, 356, 359, 368, 369, 373, 385, 390, 393]
+    foo=[202, 211, 223, 226, 227, 243, 245, 249, 258, 266, 272, 283, 304, 306, 317, 318, 356, 359, 368, 369, 373, 385, 393]
     evaluatorCombinationID=20
-    trainIDs=foo[:-6]
-    validationIDs=foo[-6:]
-#    trainIDs=foo
-#    validationIDs=[10025,10026]
+#    trainIDs=foo[:-6]
+#    validationIDs=foo[-6:]
+    trainIDs=foo
+    validationIDs=[10025,10026]
 
     evaluator_combinations=[
     [punc_loc,novelty_bool],

### 2011-05-07 paper/paper.tex
edited paper.tex

--- 2011-05-07 paper/paper.tex
@@
 
+\subsection{Walk through of one question}
+\begin{itemize}
+\item Initial setup of the Q/A system
+    \begin{itemize}
+    \item We collect all the documents that may contain the answer for all qids (20 documents per qid)
+    \item We then run the chunker on the document text creating answer candidate phrases
+    \item We save the chunker results to a set of files
+    \end{itemize}
+\item We are given a question id
+\item We retrieve the question text from the provided input file
+\item We read the chunker results 
+    \begin{itemize}
+    \item We filter out all candidate answers that are not noun phrases
+    \item We also filter out candidate answers not within 50 bytes of the aligned question in the document
+    \end{itemize}
+\item We feed the filtered answers through our answer evaluators specifically
+    \begin{itemize}
+    \item punctuation location
+    \item novelty factor
+    \item A combination of both
+    \end{itemize}
+\item The scores of the evaluators are fed into the supervised classification algorithm
+    \begin{itemize}
+    \item Spectral Regression Discriminant Analysis 
+    \item This gives us a confidence measure
+    \end{itemize}
+\item Sort by confidence measure 
+\item In descending order by confidence measure, pack them into 50-byte answers which are our final outputs
+\item The answers are then passed into our scorer
+\end{itemize}
+
 \subsection{Graphical Methods}

### 2011-05-07 paper/paper.tex
asoethusoat

@@
 the "Where" questions since there were 28 of them, which seemed like an appropriate
-amount which to train on.
+amount to train on. We ran into some prombles with some of these questions, so we
+actually only used 24 of these questions.
 
 \section{Results}
-
+\begin{table}
+\begin{tabular}{lcccc}
+%\toprule
+& \multicolumn{4}{c}{Factors} \\
+\cmidrule{2-5}
+                     & Number correct & Mean recipricol rank\\
+\cmidrule{2-5}
+Punctuation location &                &                     \\
+Novelty factor       &                &                     \\
+Both                 &                &                     \\
+\cmidrule{2-5}
+%\bottomrule
+\end{tabular}
+\caption{\label{tab:results}Results of our system on ``Where''-type
+questions. 
+\end{table}
 % we need:

### 2011-05-07 combine.py

@@ -16,7 +16,7 @@ def run_evaluators(candidates,evaluators):
 	#candidate = list of the question-candidate indexes
 	confidence = []
 	for candidate in candidates:
-#		print candidate[0]
+		print candidate[0]
 		candidateConfidence=[]
 		for evaluator in evaluators:
 			foo=evaluator(questionDict[str(candidate[4])],candidate)

### 2011-05-07 paper/paper.tex
booktabs
fix table and stuff

--- 2011-05-07 paper/paper.tex
@@
 \documentclass{article}
-
+\usepackage{booktabs}
 % required includes
@@
 to get meaningful results on which to report on. We decided to specifically analyze
-the "Where" questions since there were 28 of them, which seemed like an appropriate
-amount which to train on.
+the ``Where'' questions since there were 28 of them, which seemed like an appropriate
+amount to train on. We ran into some prombles with some of these questions, so we
+actually only used 24 of these questions.
 
 \section{Results}
-
+\begin{table}
+\begin{tabular}{lcc}
+%\toprule
+\cmidrule{2-3}
+                     & Number correct & Mean recipricol rank\\
+\cmidrule{2-3}
+Punctuation location &                &                     \\
+Novelty factor       &                &                     \\
+Both                 &                &                     \\
+\cmidrule{2-3}
+%\bottomrule
+\end{tabular}
+%\caption{\label{tab:results}Results of our system on ``Where''-type questions. 
+\end{table}
 % we need:

### 2011-05-07 one.py
changes to one

@@ -51,6 +51,7 @@ def question_candidates(q_id):
 def question_learning_data(evaluators,q_ids):
     x=[]
     y=[]
+    print q_ids
     for q_id in q_ids:
         cand=question_candidates(q_id)
         x=x+run_evaluators(cand,evaluators)

### 2011-05-07 paper/paper.tex
caption
paper

--- 2011-05-07 paper/paper.tex
@@
 \tikzstyle{process} = [rectangle, draw, text width=5em, text centered, rounded corners, minimum height=4em] 
+\tikzstyle{null} = []
 
@@
 \section{Overview}
+
+CAT-DWAM is a system that uses spectral regression discriminant analysis, a
+supervised classification algorithm, to assemble and analyze a variety of
+disparate indicators for accuracy.
+
+By using a machine learner, we aim to answer questions without relying on
+hand-written judgements of rules, although for performance reasons we were
+forced to use some rules to narrow down the input set.
 
@@
    \node[process] (ch) [below of=Q]                     {Chunker};
-   \node[data] (docs) [right of=ch]                     {Documents};
+   \node[null] (xx) [right of=ch]                     {};
+   \node[data] (docs) [right of= xx]                     {Documents};
    \node[process] (fil) [below of=ch]                     {Filter};
@@
 	\draw[->] (fil.south) to [out=270, in=90] (a3);
+	\draw[->] (Q.east) to [out=0, in=0] (fil);
 	\draw[->] (Q.west) to [out=180, in=180] (a1);
@@
 
-	We implemented a system to compare the question and answer against the
-	proposed document in two ways:  one, with a na\"ive diffing algorithm that
-	simply seeks out the longest indentical sequence, and one by implementing
-	the Smith-Waterman local alignment
+We implemented a system to compare the question and answer against the proposed
+document in two ways:  one, with a na\"ive diffing algorithm that simply seeks
+out the longest indentical sequence, and one by implementing the Smith-Waterman
+local alignment
 algorithm.\footnote{\url{http://en.wikipedia.org/wiki/Smith-Waterman_algorithm}}
-	We also implemented a system to attempt to re-write questions, discussed
-	below, and applied these same techniques to the re-written question
+We also implemented a system to attempt to re-write questions, discussed below,
+and applied these same techniques to the re-written question
 
-	The na\"ive method does an adequate job when the document highly resembles
-	the question, but does very poorly otherwise because it has no tolerance of
-	faults.  The local alignment does much better for documents that resemble
-	the question a great deal, but are not precise - it can smoothly deal with
-	small differences, and still produce a good alignment.  This is a standard
-	technique from bioinformatics, and is used extensively to align small
-	sequences of DNA, RNA or protein against a larger corpus, such as a
-	chromosome or a whole genome.  This is a fairly similar usage to aligning a
-	question against a document which is much longer than the question.
+The na\"ive method does an adequate job when the document highly resembles the
+question, but does very poorly otherwise because it has no tolerance of faults.
+The local alignment does much better for documents that resemble the question a
+great deal, but are not precise - it can smoothly deal with small differences,
+and still produce a good alignment.  This is a standard technique from
+bioinformatics, and is used extensively to align small sequences of DNA, RNA or
+protein against a larger corpus, such as a chromosome or a whole genome.  This
+is a fairly similar usage to aligning a question against a document which is
+much longer than the question.
 
-	We chose Smith-Waterman over the similar Needleman-Wunsch, of which is a
-	special case, because Needleman-Wunsch creates a \emph{global} alignment,
-	which forces all of the characters in both strings to align.  This leads to
-	the insertion of many gaps, and needlessly aligns words which are useless,
-	for example, ``Who is Queen of England?'' should align well against
-	``Elizabeth is Queen of England'' and it does, under Smith-Waterman, but the
-	``who'' causes Needleman-Wunsch to choke.  Similarly, Needleman-Wunsch does
-	not excuse one string for being shorter than the other, and chokes on all of
-	the apparent gaps that it needs to insert, so it is not a good choice.
+We chose Smith-Waterman over the similar Needleman-Wunsch, of which is a special
+case, because Needleman-Wunsch creates a \emph{global} alignment, which forces
+all of the characters in both strings to align.  This leads to the insertion of
+many gaps, and needlessly aligns words which are useless, for example, ``Who is
+Queen of England?'' should align well against ``Elizabeth is Queen of England''
+and it does, under Smith-Waterman, but the ``who'' causes Needleman-Wunsch to
+choke.  Similarly, Needleman-Wunsch does not excuse one string for being shorter
+than the other, and chokes on all of the apparent gaps that it needs to insert,
+so it is not a good choice.
 
 We locate what we best believe to be the question via these two methods, and
-	then evaluate how far away from that segment of the document the answer is.
-	If it is fairly close, that is a good indicator that it is highly related to
-	the question, but if it is far, it is likely to be unrelated.  We output
-	that distance, and either the length of the na\"ive alignment or the score
-	that Smith-Waterman produces (higher is better for both) because very short
-	or very poor alignments are less likely to reflect an actual representation
-	of the question.
+then evaluate how far away from that segment of the document the answer is.  If
+it is fairly close, that is a good indicator that it is highly related to the
+question, but if it is far, it is likely to be unrelated.  We output that
+distance, and either the length of the na\"ive alignment or the score that
+Smith-Waterman produces (higher is better for both) because very short or very
+poor alignments are less likely to reflect an actual representation of the
+question.
 
@@
 
-	1 if the candidate answer is immmediately followed by a comma, period,
-	quotation marks, semicolon, or exclamation mark, and 0 otherwise. This is a
-	good heuristic by which to rank candidate answers since it retrieves a
-	property from the context where the answer appears in the document.
+1 if the candidate answer is immmediately followed by a comma, period, quotation
+marks, semicolon, or exclamation mark, and 0 otherwise. This is a good heuristic
+by which to rank candidate answers since it retrieves a property from the
+context where the answer appears in the document.
 
@@
 
-	Because we thought it would be difficult to analyze complete parse
-	information for questions, we decided to tackle apposition by a simple
-	heuristic:  does "question, answer," align better than "question answer"?
-	If it does, then the answer is likely to be in apposition.  We also applied
-	this to re-written questions.
+Because we thought it would be difficult to analyze complete parse information
+for questions, we decided to tackle apposition by a simple heuristic:  does
+"question, answer," align better than "question answer"?  If it does, then the
+answer is likely to be in apposition.  We also applied this to re-written
+questions.
 
@@
 Answer candidates containing exclusively digits were classified as number-type
-answers, and answer candidates containing digits and other characters were classified
-as quantity-type answers.
+answers, and answer candidates containing digits and other characters were
+classified as quantity-type answers.
 
@@
 % Tom
-For quantity-type answer candidates, we used simple regular expressions to extract
-the proceeding and preceeding space-delimited phrases. This is quite close to a
-lay definition of ``word''; for the answer candidate ``about 350 km/h'', the
-phrases ``about'' and ``km/h'' would be extracted as units. We extracted the units
-from the training corpus.
+For quantity-type answer candidates, we used simple regular expressions to
+extract the proceeding and preceeding space-delimited phrases. This is quite
+close to a lay definition of ``word''; for the answer candidate ``about 350
+km/h'', the phrases ``about'' and ``km/h'' would be extracted as units. We
+extracted the units from the training corpus.
 
@@
 This evaluation measures the difference between sequence length of the question
-and the answer. It finds the length of the maximum sequence between the
-question and answer and also finds the length of the maximum sequence between
-the rewritten question and answer. A sequence is a list of (not necessarily 
+and the answer. It finds the length of the maximum sequence between the question
+and answer and also finds the length of the maximum sequence between the
+rewritten question and answer. A sequence is a list of (not necessarily
 consecutive) words that are shared by two sentences.
@@
 % Craig
-The chunker module assigns one of four part of speech tags to each answer candidate
-phrase. We discard (assign 0 to) the answer candidate if the tag is either PP (Prepositional Phrase) 
-or S (simple declarative clause). If the answer candidate is NP (Noun Phrase) we 
-assign it a probability of 1 (most likely). If the answer candidate is VP (Verb Phrase
-we assign it a probability of 0.1 (not very likely). It is possible that a higher score would be
-obtained if we also discarded verb phrases entirely, since the short question format typically
+The chunker module assigns one of four part of speech tags to each answer
+candidate phrase. We discard (assign 0 to) the answer candidate if the tag is
+either PP (Prepositional Phrase) or S (simple declarative clause). If the answer
+candidate is NP (Noun Phrase) we assign it a probability of 1 (most likely). If
+the answer candidate is VP (Verb Phrase we assign it a probability of 0.1 (not
+very likely). It is possible that a higher score would be obtained if we also
+discarded verb phrases entirely, since the short question format typically
 will point to a noun phrase.
@@
 \subsection{Score Aggregation}
-We used supervised classification algorithms to aggregate the scores produced
-by the separate answer evaluators. We trained the classifiers to predict whether
-a candidate answer was correct based on the scores produced by the answer evaluators.
+We used supervised classification algorithms to aggregate the scores produced by
+the separate answer evaluators. We trained the classifiers to predict whether a
+candidate answer was correct based on the scores produced by the answer
+evaluators.
 
 Each of the answer evaluators produced at least one numerical for each answer
-candidate; we used these as predictors for the classification algorithms.
-If we thought that some of these predictors might interact, we used their products
-as predictors as well.
+candidate; we used these as predictors for the classification algorithms.  If we
+thought that some of these predictors might interact, we used their products as
+predictors as well.
 
-For each candidate answer, we checked whether that candidate was a correct answer based on
-the regular expressions. We used the correctness of the answer (correct or incorrect)
-as the response. We were thus able to predict whether an answer was correct
-based on the scores from the answer evaluators.
+For each candidate answer, we checked whether that candidate was a correct
+answer based on the regular expressions. We used the correctness of the answer
+(correct or incorrect) as the response. We were thus able to predict whether an
+answer was correct based on the scores from the answer evaluators.
 
@@
 \end{itemize}
+
+Unfortunately, the SRDA system attempted to allocate tens of gigabytes of RAM in
+order to perform its analysis of the training data, and crashed.  So we were
+forced to thin the candidates.
 
@@
 \item We are given a question id
+    \begin{itemize}
+    \item 10025
+    \end{itemize}
 \item We retrieve the question text from the provided input file
+    \begin{itemize}
+    \item "Where is Bolivia?"
+    \end{itemize}
 \item We read the chunker results 

### 2011-05-07 paper/paper.tex

@@
 to get meaningful results on which to report on. We decided to specifically analyze
-the "Where" questions since there were 28 of them, which seemed like an appropriate
-amount which to train on.
+the ``Where'' questions since there were 28 of them, which seemed like an appropriate
+amount to train on. We ran into some prombles with some of these questions, so we
+actually only used 24 of these questions.
 
 \section{Results}
-
+\begin{table}
+\begin{tabular}{lcc}
+%\toprule
+\cmidrule{2-3}
+                     & Number correct & Mean recipricol rank\\
+\cmidrule{2-3}
+Punctuation location &                &                     \\
+Novelty factor       &                &                     \\
+Both                 &                &                     \\
+\cmidrule{2-3}
+%\bottomrule
+\end{tabular}
+\caption{\label{tab:results}Results of our system on ``Where''-type questions. }
+\end{table}
 % we need:

### 2011-05-07 paper/paper.tex
experiment design

--- 2011-05-07 paper/paper.tex
@@
 
+\section{Experiment design}
+For our experiment, we tried a few runs of the system. We ran the system with
+punctuation location on, novelty factor on and both on.
+%We repeated each of these for two training corpus sizes: 9 questions and 24 questions.
+We tested our models on two test questions and measured the number correct and
+mean recipricol rank for the test questions. Both of these measures were zero
+for all system variates (table \ref{tab:results}).
+
 \section{Results}
 \begin{table}
-\begin{tabular}{lcc}
+\begin{tabular}{llcc}
 %\toprule
-\cmidrule{2-3}
-                     & Number correct & Mean recipricol rank\\
-\cmidrule{2-3}
-Punctuation location &                &                     \\
-Novelty factor       &                &                     \\
-Both                 &                &                     \\
-\cmidrule{2-3}
+\cmidrule{2-4}
+               &       & Number correct & Mean recipricol rank\\
+\cmidrule{2-4}
+9 Training&Punctuation location &   0            &      0              \\
+questions&Novelty factor       &    0           &       0             \\
+&Both                 &             0  &                0    \\
+\cmidrule{2-4}
+24 Training&Punctuation location &                &                     \\
+questions&Novelty factor       &                &                     \\
+&Both                 &                &                     \\
+\cmidrule{2-4}
 %\bottomrule
 \end{tabular}
-%\caption{\label{tab:results}Results of our system on ``Where''-type questions. 
+\caption{\label{tab:results}Results of our system on ``Where''-type questions. }
 \end{table}

### 2011-05-07 paper/paper.tex
experiment
comment table
paper

--- 2011-05-07 paper/paper.tex
@@
    \node[process] (a3) [right of=a2]                    {Answer Evaluator};
-   \node[machine] (as) [below of=a2]                    {Score aggregator};
+   \node[machine] (as) [below of=a2]                    {Score Aggregator};
    \node[data] (A) [below of=as]                        {Answer};
@@
 document in two ways:  one, with a na\"ive diffing algorithm that simply seeks
-out the longest indentical sequence, and one by implementing the Smith-Waterman
+out the longest identical sequence, and one by implementing the Smith-Waterman
 local alignment
@@
 
-1 if the candidate answer is immmediately followed by a comma, period, quotation
+1 if the candidate answer is immediately followed by a comma, period, quotation
 marks, semicolon, or exclamation mark, and 0 otherwise. This is a good heuristic
@@
 For quantity-type answer candidates, we used simple regular expressions to
-extract the proceeding and preceeding space-delimited phrases. This is quite
+extract the proceeding and preceding space-delimited phrases. This is quite
 close to a lay definition of ``word''; for the answer candidate ``about 350
@@
 
-For test answers, we searched the context of the answer for colocated units
+For test answers, we searched the context of the answer for co-located units
 matching what we found in the training corpus. We returned a score of 1 if there
@@
 \item We used SRDA because it seemed to handle the largest sample size.
-\item We created a decision list from some of our answer evaluators. We ran the anwers
-through this decision list before sending them to the SRDA learning algorithm.
+\item We created a decision list from some of our answer evaluators. We ran the
+answers through this decision list before sending them to the SRDA learning
+algorithm.
 \end{itemize}
@@
 the ``Where'' questions since there were 28 of them, which seemed like an appropriate
-amount to train on. We ran into some prombles with some of these questions, so we
+amount to train on. We ran into some problems with some of these questions, so we
 actually only used 24 of these questions.
@@
 \begin{table}
-\begin{tabular}{lcc}
+\begin{tabular}{llcc}
 %\toprule
-\cmidrule{2-3}
-                     & Number correct & Mean recipricol rank\\
-\cmidrule{2-3}
-Punctuation location &                &                     \\
-Novelty factor       &                &                     \\
-Both                 &                &                     \\
-\cmidrule{2-3}
+\cmidrule{2-4}
+               &       & Number correct & Mean reciprocal rank\\
+\cmidrule{2-4}
+9 Training&Punctuation location &   0            &      0              \\
+questions&Novelty factor       &    0           &       0             \\
+&Both                 &             0  &                0    \\
+\cmidrule{2-4}
+24 Training&Punctuation location &                &                     \\
+questions&Novelty factor       &                &                     \\
+&Both                 &                &                     \\
+\cmidrule{2-4}
 %\bottomrule
@@
 
-\subsection{Graphical Methods}
-
-\subsection{Best Run Results}
-
 \section{Conclusion}
 
+Unfortunately, our system took much longer than expected to run, so we were
+only able to run it on very small input sets, which produced, well, awful
+results.
+
+This prevents us from drawing conclusions about the efficacy of the methods we
+used except for the following:
+
+SRDA, at least in this implementation, is probably too resource-intensive for
+use in this sort of context, where we generate tens of thousands of potential
+answer candidates and ascribe to them many different features.
+
+Either we got unlucky (with this small sample size, it is quite likely), or
+SRDA doesn't perform well with small training sets.  This is to be expected,
+but is still disheartening.
+
+Had we been able to run the system on a deeper training set and over more
+questions in the test set (and, ideally, over a validation set to tweak the
+values of various components, particularly that of the sequence aligner), we
+would have liked to make conclusions about the efficacy of the various
+techniques, particularly sequence alignment, which we believe has some promise
+that the NLP community seems to have overlooked.  Of course, it is not a very
+fast method either on these datasets, so it may also be too slow for common use.
+
 \end{document}

### 2011-05-07 paper/paper.tex

@@
 
+\section{Experiment design}
+For our experiment, we tried a few runs of the system. We ran the system with
+punctuation location on, novelty factor on and both on.
+We ran the system with a training corpus of 9 ``Where''-type questions.
+%We repeated each of these for two training corpus sizes: 9 questions and 24 questions.
+We tested our models on two test questions and measured the number correct and
+mean recipricol rank for the test questions.
+
 \section{Results}
@@
 &Both                 &             0  &                0    \\
-\cmidrule{2-4}
-24 Training&Punctuation location &                &                     \\
-questions&Novelty factor       &                &                     \\
-&Both                 &                &                     \\
+%\cmidrule{2-4}
+%24 Training&Punctuation location &                &                     \\
+%questions&Novelty factor       &                &                     \\
+%&Both                 &                &                     \\
 \cmidrule{2-4}
